---
output: github_document
title: Writing reproducible papers with RMarkdown - a repo template
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "figures/"
)
```

## User guide

This repo is designed to make reproducible and open source write-ups easy. It has minimal dependencies and is as simple as possible whilst retaining the key features that any decent paper template repo should have. These features, and how to use them, are listed below:

- [references.bib](references.bib): Add your references here in .bib format. Read more about the trusty .bib format at [bibtex.org](http://www.bibtex.org/Format/) or check-out an excellent paper on handling `.bib` files in R based on the **RefManageR** package [@mclean2014straightforward].
- [figures/](): this is where to store your figures.
- [input/](): put public facing example data here
- [output/](): folder of output data
- [README.Rmd](): the document your reading. You can write you paper as a `github_document` in the README for maximum simplicity and ease of viewing. Later you could change the output to something more fancy, e.g. using the following YAML heading, which makes use of the [**rticles**](https://github.com/rstudio/rticles) package:

```
---
title: 'The title of your paper'
author:
- Robin Lovelace (University of Leeds)
- Another Person (University of Life)
csl: ../documents/elsevier-harvard.csl # the reference format
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
fontsize: 12pt
bibliography: references.bib
---
```

## Introduction

Reproducibility is an increasingly important (some would say vital) component of academic research and, thanks to new software, is increasingly easy to implement [@jimnez2017four]. For a detailed account of how to write reproducible papers in the context of recent versions of R/RStudio, please refer to an article by [Roel Hogervorst](https://www.r-bloggers.com/writing-manuscripts-in-rstudio-easy-citations/).

For more on the subject, please see the bookdown package, which is explained in an online [book](https://bookdown.org/yihui/bookdown/) [@xie2016bookdown], and the [CRAN Task View](https://cran.r-project.org/web/views/ReproducibleResearch.html) on reproducible research.

## Methods

A great benefit of RMarkdown is that you can embed your code (revealed or hidden with `echo=FALSE`) to show exactly how your results were generated. To provide an example, below is the code that was used to generate the `references.bib` file in this repo:

```{r}
devtools::install_github("ropensci/RefManageR")
library(RefManageR)
refs = ReadPubMed(query = "Reproducibility")
pkg_ref1 = as.BibEntry(citation(package = "RefManageR"))
pkg_ref2 = as.BibEntry(citation(package = "bookdown"))
refs = c(refs, pkg_ref1, pkg_ref2)
WriteBib(refs, "references.bib")
```


```{r, echo=FALSE, eval=FALSE}
# An alternative approach to getting citations
devtools::install_github("ropensci/fulltext")
library(fulltext)
rep_citations = ft_search(query = "Reproducibility")
rep_citations$plos$data
GetBibEntryWithDOI(rep_citations$plos$data[[1]][1])
fulltext::scopus_search(query = "Reproducibility")
```

Other useful commands for setting up paper repos:

```{r}
dir.create("input") # create a folder
input1 = data.frame(x = 1:9)
input1 = readr::write_csv(input1, "input/example-input1.csv")
```


```{r, message=FALSE}
result = data.frame(y = input1 ^ 3)
readr::write_csv(x = result, "output/result.csv")
```


## Results

You can also include plot, as illustrated in the text below which generates Figure 1.

```{r basic-plot, fig.cap="A basic plot."}
plot(matrix(rpois(n = 2e3, lambda = 9), ncol = 2))
```

## Conclusion

Conclusions go here. 

Reproducible research is not only accessible, it's fun!

## References
